---
title: "Final Project"
subtitle: "Data Science for Biologists, Spring 2021"
author: "Sergio Tlatenco"
output: 
  html_document: 
    theme: readable
    highlight: haddock
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# DON'T CHANGE THE NEXT LINE, except to add NEW options if you have any.
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo = TRUE, include = TRUE)

# Load all libraries you use HERE:
library(tidyverse)

```

<br><br>

## This dataset examines tuition and cost at colleges and universities across the country. It also includes data on the diversity of US schools and post-graduation pay associated with each school.

```{r read-examine-data}
tuition_cost <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/tuition_cost.csv')

tuition_income <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/tuition_income.csv') 

salary_potential <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/salary_potential.csv')

historical_tuition <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/historical_tuition.csv')

diversity_school <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/diversity_school.csv') 

glimpse(tuition_cost)
glimpse(tuition_income)
glimpse(salary_potential)
glimpse(historical_tuition)
glimpse(diversity_school)


```

<br><br>

## Question 1

#### REPLACE TEXT WITH YOUR QUESTION. Your question should literally be a level-4 header.

### Methods

```{r question 1 code}

northeast <- c("Connecticut", "Maine", "Massachusetts", "New Hampshire", "New Jersey", "New York", "Pennsylvania", "Rhode Island", "Vermont")

southeast <- c("Alabama", "Florida", "Georgia", "Arkansas", "Louisiana", "Mississippi", "North Carolina", "South Carolina", "Tennessee")

interest_states <- c(northeast, southeast)

tuition_cost %>%
  filter(state %in% interest_states) %>%
  mutate(us_region = if_else(state %in% northeast, "Northeast", "Southeast")) %>%
  group_by(us_region, type) %>%
  summarise(mean(in_state_tuition)) 
  


```


### Answer

Replace this text with your answer to your question.


<br><br>

## Question 2

#### REPLACE TEXT WITH YOUR QUESTION. Your question should literally be a level-4 header.

### Methods

### Answer

Replace this text with your answer to your question.

<br><br>

## Question 3

#### REPLACE TEXT WITH YOUR QUESTION. Your question should literally be a level-4 header.

### Methods

### Answer

Replace this text with your answer to your question.


<br><br>

## Question 4

#### REPLACE TEXT WITH YOUR QUESTION. Your question should literally be a level-4 header.

### Methods

### Answer

Replace this text with your answer to your question.
