---
title: "Final Project"
subtitle: "Data Science for Biologists, Spring 2021"
author: "Sergio Tlatenco"
output: 
  html_document: 
    theme: readable
    highlight: haddock
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# DON'T CHANGE THE NEXT LINE, except to add NEW options if you have any.
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo = TRUE, include = TRUE)

# Load all libraries you use HERE:
library(tidyverse)

# remove exponential notation for better visual in Q2
options(scipen=999)

```

<br>

### This dataset examines tuition and cost at colleges and universities across the country. It also includes data on the diversity of US schools and post-graduation pay associated with each school.

```{r read-examine-data}

## READ IN DATA
tuition_cost <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/tuition_cost.csv')

tuition_income <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/tuition_income.csv') 

salary_potential <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/salary_potential.csv')

historical_tuition <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/historical_tuition.csv')

diversity_school <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/diversity_school.csv') 


## Preview data
glimpse(tuition_cost)
glimpse(tuition_income)
glimpse(salary_potential)
glimpse(historical_tuition)
glimpse(diversity_school)


```

<br><br>

## Question 1

#### Is private, for-profit, or public education cheaper in the Northeast or Southeast region of the U.S?

### Methods

```{r NE vs SE question 1 code}

# create variable w/ northeast states
northeast <- c("Connecticut", "Maine", "Massachusetts", "New Hampshire", "New Jersey", "New York", "Pennsylvania", "Rhode Island", "Vermont")

# create variable w/ southeast states
southeast <- c("Alabama", "Florida", "Georgia", "Arkansas", "Louisiana", "Mississippi", "North Carolina", "South Carolina", "Tennessee")

# create variable combining states from previous variables
interest_states <- c(northeast, southeast)

# start pipeline w/ appropriate data
tuition_cost %>%
  # filter for states in specified regions using saved variable combining vectors
  filter(state %in% interest_states) %>%
  # pipe mutate function to create a column that labels school by region
  mutate(us_region = if_else(state %in% northeast, "Northeast", "Southeast")) %>%
  # group schools by reigon and type of institution
  group_by(us_region, type) %>%
  # summarize by mean in state tuition under named column
  summarise(mean_in_state_tuition= mean(in_state_tuition)) %>%
  # pipe ggplot() 
  ggplot() +
  # add aes w/ x reordered and fill mapped to us_region
  aes(x = reorder(type, -mean_in_state_tuition), 
      y = mean_in_state_tuition, 
      fill = us_region) +
  # add geom_col w/ position argument
  geom_col(position = "dodge", 
           color = "gray") + 
  # add custom theme
  theme_minimal() + 
  # edit labels
  labs(x = "Type of Institution",
       y = "Mean In-State-Tuition",
       fill = "Us Region")

```


### Answer

For the three types of institutions (private, for-profit, and public), higher education is cheaper in the `Southeast` as visualized by the plot.  


<br>

## Question 2

#### How do (mean) early- and mid-career incomes compare between schools in the Northeast and Southeast regions?

### Methods

```{r NE vs SE question 2 code}

# start pipeline w/ appropriate data
salary_potential %>%
  # filter for states in specified regions using saved variable combining vectors
  filter(state_name %in% interest_states) %>%
  # pipe mutate function to create a column that labels school by region
  mutate(us_region = if_else(state_name %in% northeast, "Northeast", "Southeast")) %>%
  # group subset by region
  group_by(us_region) %>%
  # use summarise function to group by mean of a column
  summarise(mean_early_career_pay = mean(early_career_pay), 
            mean_mid_career_pay = mean(mid_career_pay)) %>%
  # use pivot_longer() to make data tidy
  pivot_longer(!us_region, names_to = "career_phase", values_to = "income") %>% 
  # pipe ggplot()
  ggplot() + 
  # add geom_point w/ color and shape mapped
  geom_point(size = 3.8, 
               aes(y = us_region, 
                   x = income, 
                   color = us_region, 
                   shape = career_phase)) +
  # add geom_segment for line between income data points
  geom_segment(aes(
      x = 56239.60, 
      y = "Northeast", 
      xend = 102302.97,
      yend = "Northeast"), 
      size = 0.1, 
      color = "red") +
  # add geom_segment for line between income data points
  geom_segment(aes(
      x = 47176.92, 
      y = "Southeast", 
      xend = 84937.18,
      yend = "Southeast"), 
      size = 0.1, 
      color = "cyan3") +
  # add custom theme
  theme_classic() +
  # edit x-axis limits for better visual
  xlim(40000, 105000) + 
  # flip order of color legend for visual purposes
  guides(color = guide_legend(reverse = TRUE)) +
  # rename shape legend
  scale_shape_discrete(labels = c("Early-Career", "Mid-Career"), 
                       name= "Career Phase") +
  # clean labels
  labs(x = "Income", 
       y = "US Region",
       color = "US Region") 
  
  

```


### Answer

The plot above shows that schools in the `Northeast` region have higher entry-level and mid-career incomes. 

<br>


## Question 3

#### Which (race) category has the highest number of enrolled students in New Jersey?

### Methods

```{r NJ  diversity question 3 code, fig.height = 5, fig.width = 9}

# rename category labels w/ line breaks
new_labels <- c("White", "Hispanic", "Black", "Asian", "Unknown", "Non-Resident\nForeign", "Two or\n More Races", "Native Hawaiian/\nPacific Islander", "American Indian/\nAlaska Native")

# pipe appropriate data
diversity_school %>%
  # filter for NJ as state
  filter(state == "New Jersey") %>%
  # remove Total Minority from category column
  filter(category != "Total Minority") %>%
  # remove Women from category column
  filter(category != "Women") %>%
  # group subset by category
  group_by(category) %>%    
  # summarize enrollment by category in new column
  summarise_at(vars(enrollment),              
               list(total_enrolled = sum)) %>%
  # pipe ggplot()
  ggplot() + 
  # add aes w/ x reordered and fill mapped to category 
  aes(x = reorder(category, -total_enrolled), y = total_enrolled, fill = category) +
  # add columns
  geom_col() +
  # add custom theme
  theme_bw() + 
  # remove legend and edit x axis text size
  theme(legend.position = "none", axis.text.x = element_text(size = 8)) + 
  # change x axis names to saved variable w/ line breaks 
  scale_x_discrete(labels = new_labels) +
  # change labels
  labs(x = "Category", 
       y = "Total Enrollment in NJ")

```



### Answer

`White` students have the highest total enrollment in NJ colleges and universities and the plot made indicates disproportionate diversity. 


<br>

## Question 4

#### Has tuition cost seen a decrease at private or public institutions since 1985?

### Methods

```{r tuition timeline question 4 code, fig.height = 5, fig.width = 9}

historical_tuition %>%
  filter(type != "All Institutions") %>%
  select(-tuition_type) %>%
  group_by(year, type) %>% 
  mutate(year = str_extract(year, "[0-9]{4}")) %>% 
  summarise_at(vars(tuition_cost),              
               list(tuition = sum)) %>%
  ggplot() +
  aes(x = year, 
      y = tuition, 
      color = type, 
      group = type) +
  geom_point() +
  geom_line() + 
  theme_grey() +
  labs(x = "Year",
       y = "Tuition",
       color = "Institution Type")


```


### Answer

Tuition costs at private and public instutions have not seen a decrease since 1985. Instead, tuition has been steadily increasing with 'Private' institutions showing a faster rate of increase in tuition cost. 
